---
title: "Create A Statistical Block"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create A Statistical Block}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This tutorial will walk you through creating a new statistical block using JavaScript and R, then writing an R function to use the block within the `gt` table. 

# Creating Blocks

To create a new statistical block for the Table Generator takes a combination of R and a little bit of JavaScript:

### Create the Drag Zone Block

Within `R/mod_TableGen_ui` you'll find a list of blocks.

If you are going to create a simple block like in the case of `MEAN`, you'll give the block an id in lowecase, and a label to be displayed in uppercase. You must also give the block `class = agg`

```{r, eval=FALSE}
tags$li(id = "mean","MEAN",class = "agg")
```

If you want to create a more complicated block that changes within the drop zone, or has hover text there are a couple more arguments you must add:

```{r, eval=FALSE}
tags$li(class = "ui-state-default agg", id = "chg",
        div(tippy(div("CHG"), "Change from Baseline")))
```

- an additional `ui-state-default` class
- another div using the package `tippy` to create the text to display `CHG` and the hover text

### Create the Drop Zone Block

Under `www/inst` you will find the `script.js` file which describes the HTML for a block within the drag zone. 

Currently there are two JavaScript functions to create HTML blocks

- `simpleBlock` which creates a block like the `FREQ` block which says "FREQ" when dragged and has a delete button

- `selectWeekBlock` which creates a dropdown when the block is dragged into the drop zone, determined by the unique values in the `AVISIT` column

You can add your block to the function `$("#droppable_agg").droppable(...` like so:

```
} else if (draggableId.includes("mean")) {
    $(this).append(selectWeekBlock(newid, "MEAN", select));
}
```

Where the lowecase `mean` corresponds to the ID of the block, and `new_id` is a JavaScript function that appends a number to your block ID in case there are multiple `mean` blocks inside the dropzone. The uppercase `MEAN` is what you want your block to display. Similarly you can change this code from `selectWeekBlock` to a simple block if you don't need the week dropdown.

### Writing the function

When you drag a column block, the shiny input returns the name of the column and the data file it came from. Using the `custom_class` function, the column name is given a class of `BDS`, `ADSL`, or `OCCDS`

We can leverage these classes when creating a new statistical function because we may want to perform slightly different calculations on an ADSL or BDS, for instance.

Let's look at `mod_tableGen_fct_freq` as an example:


### Create a function

```{r, eval=FALSE}
IDEA_freq <- function(column, week, group, data) {
  UseMethod("IDEA_freq", column)
}


IDEA_freq.default <- function(column, week, group, data) {
  rlang::abort(glue::glue(
    "Can't calculate mean because data is not classified as ADLB, BDS or OCCDS"
  ))
}
```


### Create a method for ADSL

In the case of ADSL we need to calculate the frequency for the entire table, or if groups are selected and return a table of values to be used in gt.

```{r, eval=FALSE}
IDEA_freq.ADSL <- function(column, week, group = NULL, data) {
  
  column <- rlang::sym(as.character(column))
  
  if (!is.character(data[[column]])) {
    stop(paste("Can't calculate frequency, ", column, " is not categorical"))
  }
  
  if (is.null(group)) {
    data %>%
      distinct(USUBJID, !!column) %>%
      count(!!column) %>%
      group_by(!!column) %>%
      summarise(n = sum(n)) %>%
      ungroup() %>%
      mutate(prop = n/sum(n)) %>%
      mutate(x = paste0(n, " (", round(prop*100, 2), ")")) %>%
      select(!!column, x)
  } else {
    
    if (group == column) {
      stop(glue::glue("Cannot calculate frequency for {column} when also set as group."))
    }
    
    group <- rlang::sym(group)
    data %>%
      distinct(USUBJID, !!column, !!group) %>%
      count(!!column, !!group) %>%
      group_by(!!group) %>%
      mutate(prop = prop.table(n)) %>%
      mutate(v1 = paste0(n, ' (', round(prop*100, 2), ')')) %>%
      select(-n, -prop) %>% 
      spread(!!group, v1)
  }
}
```

### Create a method for BDS

We don't need a method for BDS because we currently only import PARAMCD blocks and since AVAL and CHG are numeric we can't calculate frequency:

```{r, eval=FALSE}
IDEA_freq.BDS <- function(column, week, group = NULL, data) {
  rlang::abort(glue::glue(
    "Can't calculate frequency for BDS - {column} is numeric"
  ))
}
```

Feel free to explore the mean, ANOVA, and chg files for other statistical examples. For instance, blocks with BDS methods use the `week` argument which is supplied from the block's dropdown. 

# Putting it all together

Once you have your statistical function, you need to add it to the `IDEA_methods` function:

```{r, eval=FALSE}
IDEA_methods <- function(agg, column, week, group, data) {
  if (agg == "MEAN") {
    IDEA_mean(column, week, group, data)
  } else if (agg == "FREQ") {
    IDEA_freq(column, week, group, data)
  } else if (agg == "ANOVA") {
    IDEA_anova(column, week, group, data)
  } else {
    IDEA_chg(column, week, group, data)
  }
}
```

This function will look for the name of the dragged in statistical block and apply the correct statistical function based on the block's name. Since the first argument of the function is the column to apply the statistical block on, it will look for the data file the column came from and choose the correct method. 

Finally this function is used within a `map` in `mod_tableGen` so each statistical block is applied to each column in the drop zone iteratively. 